Using evolutionary algorithms to optimize the estimation of a diffusion process\titlenote{(Produces the permission block, and copyright information). For use with SIG-ALTERNATE.CLS. Supported by ACM.}

### 2015-02-07 paper.tex
paperGecco.tex con el formato terminado ahora el texto
Revisando la nueva introducción
@mgarenas he visto una palabra que no existe, "transity". ¿Puedes revisarla, por favor? El resto está correctamente escrito, pero sería conveniente que alguien, @pacastillo o @nuriastatgirl lo revisaran.

\documentclass[runningheads,a4paper]{llncs}

@@

\begin{document}
\mainmatter  % start of an individual contribution

@@

\begin{abstract}
The maximum likelihood estimation of the parameters of a diffusion
process requires a system of equations that, for some cases, has no
explicit solution; thus, the solution has to be estimated. We will use
optimization methods to find those solutions: an Iterative Method, an
algorithm based on Newton-Raphson solver, a Variable Neighbourhood
Search method, a Simulated Annealing algorithm and an Evolutionary
Algorithm. We generate four data sets following a Gompertz-lognormal
diffusion process using different noise levels. The methods are applied
to these data sets in order to estimate the parameters that rule
the diffusion process. Results show that bio-inspired methods
yield suitable solutions for the problem every time, even when the
noise level increases. On the other hand, some analytical methods as
Newton-Raphson or the Iterative Method do not always solve the
problem whether their scores depend on the starting point for initial
solution or the noise level hinders the resolution of the problem. In
these cases, the bio-inspired algorithms remain as a suitable and
reliable approach.

>>>>

\documentclass{sig-alternate}
%\documentclass[runningheads,a4paper]{llncs}

@@

\begin{document}
 \conferenceinfo{GECCO'15,} {July 11-15, 2015, Madrid, Spain.}
    \CopyrightYear{2015}
    \crdata{TBA}
    \clubpenalty=10000
    \widowpenalty = 10000
\mainmatter  % start of an individual contribution

@@

\begin{abstract}
The estimation of the  maximum likelihood of the parameters of a diffusion
process requires to solve a system of equations that, for some cases, has no
explicit solution; thus, it has to be estimated. We will use
optimization methods to find those solutions: an Iterative Method, an
algorithm based on Newton-Raphson solver, a Variable Neighbourhood
Search method, a Simulated Annealing algorithm and an Evolutionary
Algorithm. We generate four data sets following a Gompertz-lognormal
diffusion process using different noise levels. The methods are applied
to these data sets in order to estimate the parameters that rule
the diffusion process. Results show that bio-inspired methods
yield suitable solutions for the problem every time, even when the
noise level increases. On the other hand, some analytical methods as
Newton-Raphson or the Iterative Method do not always solve the
problem whether their scores depend on the starting point for initial
solution or the noise level hinders the resolution of the problem. In
these cases, the bio-inspired algorithms remain as a suitable and
reliable approach.

### 2015-02-07 paperGecco.tex
Revisando la nueva introducción
@mgarenas he visto una palabra que no existe, "transity". ¿Puedes revisarla, por favor? El resto está correctamente escrito, pero sería conveniente que alguien, @pacastillo o @nuriastatgirl lo revisaran.

\numberofauthors{6} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
%Nuria Rico\titlenote{Dr.~Nuria Rico Castro.}\\
%       \affaddr{Faculty of Science }\\
%       \affaddr{University of Granada}\\
%       \email{nrico@ugr.es}       
Nuria Rico\titlenote{Dr.~Rico}\\
       \affaddr{Statistics and Operational Research}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{nrico@ugr.es}
% 2nd. author
\alignauthor
M. G. Arenas\titlenote{Dr. ~Arenas}\\
       \affaddr{Computer Architecture}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{mgarenas@ugr.es}
% 3rd. author
\alignauthor 
Desir\'{e}e Romero\titlenote{Dr. ~Romero}\\
       \affaddr{Statistics and Operational Research}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{deromero@ugr.es}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor 
P. A. Castillo\titlenote{Dr. ~Castillo}\\
       \affaddr{Computer Architecture}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{pacv@ugr.es}
% 5th. author
\alignauthor
J. M. Crespo\titlenote{Dr. ~Crespo}\\
       \affaddr{Statistics and Operational Research}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{crespo@ugr.es}
% 6th. author
\alignauthor
J. J. Merelo\titlenote{Dr. ~Merelo}\\
       \affaddr{Computer Arquitecture}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{jmerelo@ugr.es}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{1 Feb 2015}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

\begin{abstract}
The maximum likelihood estimation of the parameters of a diffusion
process requires a system of equations that, for some cases, has no
explicit solution; thus, the solution has to be estimated. We will use
optimization methods to find those solutions: an Iterative Method, an
algorithm based on Newton-Raphson solver, a Variable Neighbourhood
Search method, a Simulated Annealing algorithm and an Evolutionary
Algorithm. We generate four data sets following a Gompertz-lognormal
diffusion process using different noise levels. The methods are applied
to these data sets in order to estimate the parameters that rule
the diffusion process. Results show that bio-inspired methods
yield suitable solutions for the problem every time, even when the
noise level increases. On the other hand, some analytical methods as
Newton-Raphson or the Iterative Method do not always solve the
problem whether their scores depend on the starting point for initial
solution or the noise level hinders the resolution of the problem. In
these cases, the bio-inspired algorithms remain as a suitable and
reliable approach.

@@

Diffusion process is one usual mathematical formulation of a model for population growth in a randomly
varying environment \cite{lognormal}. These models describe how a population adopts a new innovation, technology or behaviour. With these kind of models, we have the transity density function (t.d.f.) which express the chance of observing
any size for the population at any time when the initial size is specified. With the t.d.f we also are able to determine the time at that population first reaches a fixed size. Moreover, we can forecast future values of the population size as well as confidence intervals for the predictions.

In this paper we focus our attention on the Gompertz-lognormal diffusion process \cite{proceedingBol}. It can be seen as a mixture between a Gompertz-type process \cite{Gut07b} and a lognormal process \cite{GutCy}. The first one models precesses with trends showing enclosed growth patterns, as biological growth with finite space or resources. The second one is useful for modelling not bounded growth with exponential trend.

The mixture of them, the Gompertz-lognormal diffusion process, combines both trends; a first part Gompertz-like growth and a second exponential part. This process is defined through the infinitesimal moments
%{\scriptsize
\[
\begin{array}{l}
A_1(x,t)=(m e^{-\beta t}+c)x \\
A_2(x,t)=\sigma^2 x^2,
\end{array}
\]
with $m > \beta > 0$, $c\in \mathbb{R}$.
%}
%Maribel: Deber{\'\i}amos poner qu{\'e} es m, que es e, que es c y por supuesto mencionar que t es el tiempo y que x es la variable independiente no?

>>>>

\numberofauthors{6} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
%Nuria Rico\titlenote{Dr.~Nuria Rico Castro.}\\
%       \affaddr{Faculty of Science }\\
%       \affaddr{University of Granada}\\
%       \email{nrico@ugr.es}       
Nuria Rico\titlenote{Dr.~Rico}\\
       \affaddr{Statistics and Operational Research}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{nrico@ugr.es}
% 2nd. author
\alignauthor
M. G. Arenas\titlenote{Dr. ~Arenas}\\
       \affaddr{Computer Architecture}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{mgarenas@ugr.es}
% 3rd. author
\alignauthor 
Desir\'{e}e Romero\titlenote{Dr. ~Romero}\\
       \affaddr{Statistics and Operational Research}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{deromero@ugr.es}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor 
P. A. Castillo\titlenote{Dr. ~Castillo}\\
       \affaddr{Computer Architecture}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{pacv@ugr.es}
% 5th. author
\alignauthor
J. M. Crespo\titlenote{Dr. ~Crespo}\\
       \affaddr{Statistics and Operational Research}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{crespo@ugr.es}
% 6th. author
\alignauthor
J. J. Merelo\titlenote{Dr. ~Merelo}\\
       \affaddr{Computer Architecture}\\
       \affaddr{University of Granada}\\
       \affaddr{Granada, Spain}\\
       \email{jmerelo@ugr.es}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{1 Feb 2015}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

\begin{abstract}
The maximum likelihood estimation of the parameters of a diffusion
process requires to solve a system of equations that, in some cases, has no
explicit solution; thus, it has to be estimated. We will use
optimization methods to find those solutions: an Iterative Method, an
algorithm based on Newton-Raphson solver, a Variable Neighbourhood
Search method, a Simulated Annealing algorithm and an Evolutionary
Algorithm. We generate four data sets following a Gompertz-lognormal
diffusion process using different noise levels. The methods are applied
to these data sets in order to estimate the parameters that rule
the diffusion process. Results show that bio-inspired methods
yield suitable solutions for the problem every time, even when the
noise level increases. On the other hand, some analytical methods as
Newton-Raphson or the Iterative Method do not always solve the
problem whether their scores depend on the starting point for initial
solution or the noise level hinders the resolution of the problem. In
these cases, the bio-inspired algorithms remain as a suitable and
reliable approach.

@@

A diffusion process is an usual mathematical formulation of a model for population growth in a randomly
varying environment \cite{lognormal}. These models describe how a
population adopts a new innovation, technology or behaviour. With
these kind of models, we have the transity density function (t.d.f.)
% transity? Is that correct? - JJ
which express the chance of observing 
any size for the population at any time when the initial size is
specified. With the t.d.f we also are able to determine the time at
that population first reaches a fixed size. Moreover, we can forecast
future values of the population size as well as confidence intervals
for the predictions.

In this paper we focus our attention on the Gompertz-lognormal
diffusion process \cite{proceedingBol}. It can be seen as a mixture
between a Gompertz-type process \cite{Gut07b} and a lognormal process
\cite{GutCy}. The first one models precesses with trends showing
enclosed growth patterns, as biological growth with finite space or
resources. The second one is useful for modelling unbounded growth
with an exponential tendency.

The mixture of them, the Gompertz-lognormal diffusion process,
combines both trends; a first part Gompertz-like growth and a second
exponential part. This process is defined through the infinitesimal
moments 
%{\scriptsize
\[
\begin{array}{l}
A_1(x,t)=(m e^{-\beta t}+c)x \\
A_2(x,t)=\sigma^2 x^2,
\end{array}
\]
with $m > \beta > 0$, $c\in \mathbb{R}$.
%}
%Maribel: Deber{\'\i}amos poner qu{\'e} es m, que es e, que es c y por
%supuesto mencionar que t es el tiempo y que x es la variable
%independiente no?
